### 2016-05-02 ch03-lrm/truncation.tex
Most comments of Johan on CH03

The main contribution of this section is to come up with an improved \gls{FRF} by developing a method for smoothing the \gls{FRF} estimated via the \gls{LPM}. 
This is done by truncating the associated impulse response, as in~\citet{Schoukens1998}, but with the following additional extension: the determination of the optimal truncation time, without any user interaction, in conjunction with the use of the \gls{LPM} for leakage reduction. 
The truncation point (cut-off index) is determined by one of two statistically inspired methods.
It is then shown that a smooth \gls{LPM} estimate of the \gls{FRF} lowers the variance, thus improving the assessment level of the dynamic properties of a given \gls{LTI} system.

@@

\subsection*{Theory on the LPM and Windowing}
\label{sec:nparam:trunc:theoryLPMandWindowing}

Details of the theory pertaining to this section may be found in \citet{Schoukens2009LPM}, basic aspects of which are presented below.

\subsubsection{Spectral Nonparametric Model and Assumptions}

Using convolution, the noiseless output signal in \figref{fig:nparam:blockH0} is $y_0(t) = g_0(t)*u_0(t)$, where the signals are sampled at $t = 0, 1, 2,...,N-1$ and the excitation signal $u_0(t)$ is random noise. 
As shown in \citet{Pintelon2012} and \citet{Pintelon1997ARB} the \gls{DFT} spectra of the sampled signals at the frequency $\omega_k$ are related as follows:
\begin{equation}\label{eq:nparam:trunc:LPM:leakage}
Y_0(k)=G_0(\omega_k)U_0(k)+T_G(\omega_k)
\end{equation}
where $U_0(k)$ is the \gls{DFT} spectrum of the random noise excitation,  $G_0(\omega_k)$ is the transfer function and $T_G(\omega_k)$ is the transients leakage term.
In this section, the \gls{DFT} of a signal $x(t)$ is defined explicitly as~\citep{Oppenheim1983}

\begin{equation}\label{eq:defDFT}
X(\omega_k) = \frac{1}{\sqrt{N}}\sum_{t=0}^{N-1}x(t)e^{-\frac{j2\pi kt}{N}}
\end{equation}

The following assumptions are consistent with \citep{Schoukens2009LPM} for an \gls{OE} setup (\figref{fig:nparam:blockH0}):
\begin{assumption}
The \gls{SISO} system $G_0$ and  the transients leakage spectrum $T_G$ are both smooth functions of the frequency.
\end{assumption}

\begin{assumption}
The spectrum of the  input signal $U_0$ is not a smooth function of the frequency, but a rough function, i.e. $U_0(k+1) - U_0(k)$ should not vanish to zero~\citep{Schoukens2009LPM}.
\end{assumption}

In other words, for the excitation signal to be rough: the magnitude of the spectral difference $|U_0(k+1) - U_0(k)|$ must have a probability of 1 (unity) for it to remain in the same order of magnitude as $|U_0(k)|$, irrespective of the record length $N\rightarrow\infty$ and the corresponding  frequency resolution $f_0\rightarrow{0}$.

\begin{assumption}
The measured output signal is $y(t) = y_0(t) + v(t)$ with the exact input signal $u_0(t)$ known. 
The filtered white noise $v(t) = H_0(q)e(t)$ is the disturbing noise source at the output.
\end{assumption}

\subsubsection{The Hanning Window and the \glsentrytext{LPM}}
\label{sec:nparam:trunc:LPMFRFest}
\paragraph{Windowing}
The effect of windowing may be  exemplified by use of the Hanning window, which is a popular choice.
In a nutshell, the Hanning window is effective in reducing  the leakage errors, but  with a trade off of increased interpolation errors. 
Pertinent details of the Hanning window are discussed in \citep{Schoukens2006LPM,Antoni2007FRF,Schoukens2009LPM,Wellstead1981} and \citep{Harris1978}.

The anomalies due to interpolation errors, associated with windowing, can easily be mitigated or circumvented by use of the \gls{LPM}. 
As shown in \citep{Pintelon2012}, the leakage errors are reduced from  $\bigO{{N}^{-1}}$, for Hanning windowing, to $\bigO{{N}^{-3}}$ or better, for the \gls{LPM}.

@@

The estimation of $\hat{G}_{\LPM}(\omega_k)$ at a single frequency $\omega_k = 0.586\unit{Hz}$ is illustrated in \figref{LPM_Schematic_EG}.% centered at $k$, where $r = 0$.
\begin{figure}[htb] %  figure placement: here, top, bottom, or page
   \centering
   \setlength{\figurewidth}{0.8\columnwidth}
   \setlength{\figureheight}{0.68\figurewidth}
   \input{\thisDir/figs/LPMfig.tikz}
   \caption[Illustration of the LPM.]{Illustration of the \gls{LPM}. Top figures: input spectrum (left) and output spectrum (right). 
   A small band is selected (gray area), in which $\hat G_{\LPM}(\omega_{k+r})$ is estimated (black thick line, bottom figure) as a local polynomial. Only its central value $\hat G_{\LPM}(\omega_k)$ (white circle) is retained. This procedure is repeated for all frequencies $\omega_k$.}
   \label{LPM_Schematic_EG}
\end{figure}
This procedure is repeated at each central frequency $\omega_k$ for  $k\in \Set{1,\dots,N/2}$ (i.e. the entire frequency band, including Nyquist, but excluding \gls{DC}, and regularizing at the boundaries). This completes the \gls{LPM}, yielding an intrinsic averaging of the spectra of the estimated \gls{FRF}, $\hat{G}_{\LPM}(\omega_k)$.

@@

\begin{assumption}\label{ass:imprespdecay}
The impulse response $g(t)$ decays exponentially over time, i.e. $\exists A \in \PositiveRealNumbers, \exists a \in \NegativeRealNumbers \without \Set{0}, \forall t \in \PositiveRealNumbers: \abs{g(t)} \leq A \exp(- a t)$.
\end{assumption}

@@

\subsubsection{Truncation Via the $F$-Test}
\label{sec:nonparametric:truncation:ftest}
To distinguish the part of the measured impulse response where the actual impulse response dominates from the part where the noise dominates, the estimated impulse response $\hat{g}_{\LPM}$ is split into smaller segments. 
One can then gradually determine whether these segments are noise or the actual impulse response.

>>>>

The main contribution of this section is to come up with an improved \gls{FRF} by developing a method for smoothing the \gls{FRF} estimated via the \gls{LPM}. 
This is done by truncating the associated impulse response, as in~\citet{Schoukens1998}, but with the following additional extension: the determination of the optimal truncation time, without any user interaction, in conjunction with the use of the \gls{LPM} for leakage reduction. 
The truncation point (cut-off index) is determined by one of two statistically inspired methods.
It is then shown that a smooth \gls{LPM} estimate of the \gls{FRF} lowers the variance, thus improving the assessment level of the dynamic properties of a given \gls{LTI} system.
It should be noted, however, that such smoothing is merely of importance when inspecting the \gls{FRF} visually.
In particular, a slight bias is incurred when smoothing the \gls{FRF}.
If one is to fit a parametric model to the smoothed \gls{FRF}, one essentially ends up with a biased model (which is often undesirable).

@@

% \subsection*{Theory on the LPM and Windowing}
% \label{sec:nparam:trunc:theoryLPMandWindowing}

% Details of the theory pertaining to this section may be found in \citet{Schoukens2009LPM}, basic aspects of which are presented below.

% \subsubsection{Spectral Nonparametric Model and Assumptions}

% Using convolution, the noiseless output signal in \figref{fig:nparam:blockH0} is $y_0(t) = g_0(t)*u_0(t)$, where the signals are sampled at $t = 0, 1, 2,...,N-1$ and the excitation signal $u_0(t)$ is random noise. 
% As shown in \citet{Pintelon2012} and \citet{Pintelon1997ARB} the \gls{DFT} spectra of the sampled signals at the frequency $\omega_k$ are related as follows:
% \begin{equation}\label{eq:nparam:trunc:LPM:leakage}
% Y_0(k)=G_0(\omega_k)U_0(k)+T_G(\omega_k)
% \end{equation}
% where $U_0(k)$ is the \gls{DFT} spectrum of the random noise excitation,  $G_0(\omega_k)$ is the transfer function and $T_G(\omega_k)$ is the transients leakage term.
% In this section, the \gls{DFT} of a signal $x(t)$ is defined explicitly as~\citep{Oppenheim1983}

% \begin{equation}\label{eq:defDFT}
% X(\omega_k) = \frac{1}{\sqrt{N}}\sum_{t=0}^{N-1}x(t)e^{-\frac{j2\pi kt}{N}}
% \end{equation}

% The following assumptions are consistent with \citep{Schoukens2009LPM} for an \gls{OE} setup (\figref{fig:nparam:blockH0}):
% \begin{assumption}
% The \gls{SISO} system $G_0$ and  the transients leakage spectrum $T_G$ are both smooth functions of the frequency.
% \end{assumption}

% \begin{assumption}
% The spectrum of the  input signal $U_0$ is not a smooth function of the frequency, but a rough function, i.e. $U_0(k+1) - U_0(k)$ should not vanish to zero~\citep{Schoukens2009LPM}.
% \end{assumption}

% In other words, for the excitation signal to be rough: the magnitude of the spectral difference $|U_0(k+1) - U_0(k)|$ must have a probability of 1 (unity) for it to remain in the same order of magnitude as $|U_0(k)|$, irrespective of the record length $N\rightarrow\infty$ and the corresponding  frequency resolution $f_0\rightarrow{0}$.

% \begin{assumption}
% The measured output signal is $y(t) = y_0(t) + v(t)$ with the exact input signal $u_0(t)$ known. 
% The filtered white noise $v(t) = H_0(q)e(t)$ is the disturbing noise source at the output.
% \end{assumption}

% \subsubsection{The Hanning Window and the \glsentrytext{LPM}}
% \label{sec:nparam:trunc:LPMFRFest}
% \paragraph{Windowing}
% The effect of windowing may be  exemplified by use of the Hanning window, which is a popular choice.
% In a nutshell, the Hanning window is effective in reducing  the leakage errors, but  with a trade off of increased interpolation errors. 
% Pertinent details of the Hanning window are discussed in \citep{Schoukens2006LPM,Antoni2007FRF,Schoukens2009LPM,Wellstead1981} and \citep{Harris1978}.

% The anomalies due to interpolation errors, associated with windowing, can easily be mitigated or circumvented by use of the \gls{LPM}. 
% As shown in \citep{Pintelon2012}, the leakage errors are reduced from  $\bigO{{N}^{-1}}$, for Hanning windowing, to $\bigO{{N}^{-3}}$ or better, for the \gls{LPM}.

@@

The estimation of $\hat{G}_{\LPM}(\omega_k)$ at a single frequency $\omega_k = 0.586\unit{Hz}$ is illustrated in \figref{LPM_Schematic_EG}.% centered at $k$, where $r = 0$.
\begin{figure}[htb] %  figure placement: here, top, bottom, or page
   \centering
   \setlength{\figurewidth}{0.8\columnwidth}
   \setlength{\figureheight}{0.68\figurewidth}
   \input{\thisDir/figs/LPMfig.tikz}
   \caption[Illustration of local modeling.]{Illustration of local modeling. Top figures: input spectrum (left) and output spectrum (right). 
   A small band is selected (gray area), in which $\hat G_{\LPM}(\omega_{k+r})$ is estimated (black thick line, bottom figure) as a local polynomial. Only its central value $\hat G_{\LPM}(\omega_k)$ (white circle) is retained. This procedure is repeated for all frequencies $\omega_k$. \disclaimer{Figure based on \citep[Fig.~2]{Lumori2014TIM}.}}
   \label{LPM_Schematic_EG}
\end{figure}
This procedure is repeated at each central frequency $\omega_k$ for  $k\in \Set{1,\dots,N/2}$ (i.e. the entire frequency band, including Nyquist, but excluding \gls{DC}, and regularizing at the boundaries). This completes the \gls{LPM}, yielding an intrinsic averaging of the spectra of the estimated \gls{FRF}, $\hat{G}_{\LPM}(\omega_k)$.

@@

\begin{assumption}\label{ass:imprespdecay}
The impulse response $g(t)$ decays exponentially over time, i.e. 
\[
  \exists A \in \PositiveRealNumbers, 
  \exists a > 0 ,
  \exists t_0 > 0,
  \forall t > t_0: 
     \abs{g(t)} \leq A \exp(- a t) \text{.}
\]
\end{assumption}

@@

\subsubsection{Truncation Via the $F$-Test}
\label{sec:nonparametric:truncation:ftest}
To distinguish the part of the measured impulse response where the actual impulse response dominates from the part where the noise dominates, the estimated impulse response $\hat{g}_{\LPM}$ is split into smaller segments. 
One can then gradually determine if in each of these segments, the observed impulse response is dominated by the noise.
